---
title: "StatisticAnalysis"
output: html_document
---

#### Satisfaction Level

```{r fig.width=10, warning=FALSE}
# Plot satisfaction level using histogram and density plot. 

p1 <- ggplot(hr_stat, aes(Satisfaction, fill = Quit, colour = Quit)) + 
  geom_histogram(position = "fill", binwidth = 0.01, alpha = 0.8) +
  ggtitle("Satisfaction Ratings of Employees Histogram") +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major.y = element_line(colour = "grey80"),
        legend.position = "none") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), labels = scales::percent) +
  labs(y = "Percentage") +
  scale_fill_discrete(name = "Stay/Quity", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit"))

p2 <- ggplot(hr_stat, aes(Satisfaction, fill = Quit, colour = Quit)) + 
  geom_density(position = "identity", binwidth = 0.01, alpha = 0.6) +
  ggtitle("Satisfaction Ratings of Employees Density Plot") + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_x_continuous(breaks = seq(0,1,0.1)) +
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit"))

grid.arrange(p1, p2, ncol = 2)

```


* The plot indicates that most employees who left have a low satisfaction level between 0.37 - 0.50. 
* There is a tri-modal effect. Satisfaction levels of (< 15), (0.35 - 0.50), (0.7-0.9) left the company more.
* From the individuals who stayed, we can see a general trend of having 50% or higher satisfaction. 

#### Last Evaluation 

```{r fig.width=10, warning=FALSE}


p3 <- ggplot(hr_stat, aes(Evaluation, fill = Quit, colour = Quit)) +
  geom_histogram(position = "fill", binwidth = 0.01, alpha = 0.8) +
  ggtitle("Last Evaluation of Employees Histogram") + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major.y = element_line(colour = "grey80"), 
        legend.position = "none") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), labels = scales::percent) +
  labs(y = "Percentage") +
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit"))

p4 <- ggplot(hr_stat, aes(Evaluation, fill = Quit, colour = Quit)) + 
  geom_density(position = "identity", binwidth = 0.01, alpha = 0.6) + 
  ggtitle("Last Evaluation of Employees Density Plot") + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit"))

grid.arrange(p3, p4, ncol = 2)
```

* Bi-modal relationship between quitting and company evaluation. 
* The company is losing many of their top evaluated performers. 
* Individuals who are staying have an evaluation of above 40%.

#### Number of Projects 
```{r}

position <- data.frame(pos = c(.80, .07, .80, .07, .80, .07, .80, .07, .80, .07, .07))
testing <- data.frame(projects = hr_stat[,3], Quit = hr_stat[,7])
testingproject <- testing %>%
  group_by(projects, Quit) %>%
  summarise(count = n()) %>%
  mutate(pct = count/sum(count)) %>% 
  mutate(pcttot = count/14999*100) 
testingproject$pcttot <- round(testingproject$pcttot, digits = 2)

ggplot(testingproject, aes(projects, pct, colour = Quit, fill = Quit)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  geom_text(data = testingproject, 
            aes(x = projects, y = position$pos, label = paste0(pcttot, "%")),
            colour = "black", size = 3) +
  geom_text(data = testingproject, 
            aes(x = projects, y = position$pos + 0.06, label = paste0(count)),
            colour = "black", size = 3) +
  ggtitle("Number of Projects, Stay vs Quit") + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major.y = element_line(colour = "grey80")) +
  scale_x_continuous(breaks = seq(2, 7, 1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.10), labels = scales::percent) +
  labs(y = "Relative Percentage", x = "Number of Projects",
       subtitle = "Bars note number of employees and percentage of whole company.") + 
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) 

```

* Individuals with 2, (4-6+) projects are more likely to quit.
* 65% of individuals with 2 and 6+ projects have quit, 16.53% of the whole company. 
* Most employees with 3-4 projects have stayed with the company, 52.94% of employees. 
* Employees with 2 projects, 65% have quit, which is about half of total employee who quit. 

#### Average Montly Hours Worked

```{r fig.width=10, warning=FALSE}
p5 <- ggplot(hr_stat, aes(AvgMonthlyHours, fill = Quit, colour = Quit)) + 
  geom_histogram(binwidth = 1, position = "fill", alpha = 0.8) + 
  ggtitle("Average Monthly Hours Histogram") + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major.y = element_line(colour = "grey80"),
        legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 1, 0.1), labels = scales::percent) +
  labs(y = "Percentage") +
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit"))

p6 <- ggplot(hr_stat, aes(AvgMonthlyHours, fill = Quit, colour = Quit)) +
  geom_density(position = "identity", binwidth = 1, alpha = 0.6) +
  ggtitle("Average Monthly Hours Density Plot") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom") +
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit"))

grid.arrange(p5, p6, ncol = 2)
```

* Bi-modal relationship, many employees who left either worked under 175 hours or above 225. 
* There is a higher percentage of employees quitting with 250+ hours. 
* Employees who are underworked and overworked are quitting. 

#### Time spent with company

```{r}
Years <- data.frame(pos = c(.80, .07, .80, .07, .80, .07, .80, .07, .80, .07, .8, .8, .8))
testing <- data.frame(years = hr_stat[,5], Quit = hr_stat[,7])
testingyears <- testing %>%
  group_by(years, Quit) %>%
  summarise(count = n()) %>%
  mutate(pct = count/sum(count)) %>% 
  mutate(pcttot = count/14999*100) 
testingyears$pcttot <- round(testingyears$pcttot, digits = 2)

ggplot(testingyears, aes(years, pct, colour = Quit, fill = Quit)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  geom_text(data = testingyears, 
            aes(x = years, y = Years$pos, label = paste0(pcttot, "%")),
            colour = "black", size = 3) +
  geom_text(data = testingyears, 
            aes(x = years, y = Years$pos + 0.06, label = paste0(count)),
            colour = "black", size = 3) +
  ggtitle("Years in Company") + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major.y = element_line(colour = "grey80")) +
  scale_x_continuous(breaks = seq(1, 10, 1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.10), labels = scales::percent) +
  labs(y = "Relative Percentage", x = "Years",
       subtitle = "Bars note number of employees and percentage of whole company.") + 
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) 

```

* Most employees left between working 3-6 years with the company, 23.44% of the company.
* Of the employees who have been with the company for 5 years, 50% have quit. 


#### Promotions


```{r}
position <- data.frame(posy = c(.75, .10, .75, .10),
                       posx = c(1.08, 1.08, 2.08, 2.08))
testing <- data.frame(promotion = hr_stat[,8], Quit = hr_stat[,7])
testingpromotion <- testing %>%
  group_by(promotion, Quit) %>%
  summarise(count = n()) %>%
  mutate(pct = count/sum(count)) %>% 
  mutate(pcttot = count/14999*100) 
testingpromotion$pcttot <- round(testingpromotion$pcttot, digits = 4)

ggplot(testingpromotion, aes(promotion, pct, colour = Quit, fill = Quit)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  geom_text(data = testingpromotion, 
            aes(x = promotion, y = position$posy, label = paste0(pcttot, "%")),
            colour = "black", size = 3) +
  geom_text(data = testingpromotion, 
            aes(x = position$posx, y = position$posy, label = paste0(count)),
            colour = "black", size = 3) +
  ggtitle("Promotions of Employees") + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, angle = 90),
        panel.grid.major.x = element_line(colour = "grey80")) +
  scale_x_discrete(labels = c("1" = "Promotion", "0" = "No Promotion")) +
  scale_y_continuous(breaks = seq(0, 1.0, 0.1), labels = scales::percent) +
  labs(y = "Relative Percentage", x = "Promotions",
       subtitle = "Bars note number of employees and percentage of whole company.")+ 
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  coord_flip()

```

* From this table we see that most individuals who left were not offered a promotion in the last 5 years. 
* 25% of employees who were not offered a promotion have quit. 
* 5% of promoted employees have quit. 

#### Salary

```{r}

position <- data.frame(posy = c(.75, .07, .75, .07, .75, .07),
                       posx = c(1.1, 1.1, 2.1, 2.1, 3.1, 3.1))
testing <- data.frame(salary = hr_stat[,10], Quit = hr_stat[,7])
testingsalary <- testing %>%
  group_by(salary, Quit) %>%
  summarise(count = n()) %>%
  mutate(pct = count/sum(count)) %>% 
  mutate(pcttot = count/14999*100) 
testingsalary$pcttot <- round(testingsalary$pcttot, digits = 2)

ggplot(testingsalary, aes(salary, pct, colour = Quit, fill = Quit)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  geom_text(data = testingsalary, 
            aes(x = salary, y = position$posy, label = paste0(pcttot, "%")),
            colour = "black", size = 3) +
  geom_text(data = testingsalary, 
            aes(x = position$posx, y = position$posy, label = paste0(count)),
            colour = "black", size = 3) +
  ggtitle("Salary Level of Employees") + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, angle = 90),
        panel.grid.major.x = element_line(colour = "grey80")) +
  scale_x_discrete() +
  scale_y_continuous(breaks = seq(0, 1.0, 0.1), labels = scales::percent) +
  labs(y = "Relative Percentage", x = "Salary Level",
       subtitle = "Bars note number of employees and percentage of whole company.")+ 
  scale_fill_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  scale_color_discrete(name = "Stay/Quit", labels = c("0" = "Stay", "1" = "Quit")) +
  coord_flip()

  
```

* Most of the employees who quit were in the low and medium bracket of salary. 
* About 6.6% of high paid employees have quit. 
* 20% of medium paid employees have quit.
* 30% of low paid employees have quit. 